Prediction of US election with linear model

### 2024-10-21 paper/paper.qmd
update data and measurement

library(tidyverse)
library(palmerpenguins)
```

@@

We use the statistical programming language R [@citeR].... Our data [@shelter].... Following @tellingstories, we consider...

Overview text

@@

Some paragraphs about how we go from a phenomena in the world to an entry in the dataset.

## Outcome variables

Add graphs, tables and text. Use sub-sub-headings for each outcome variable or update the subheading to be singular.

Some of our data is of penguins (@fig-bills), from @palmerpenguins.

```{r}
#| label: fig-bills
#| fig-cap: Bills of penguins
#| echo: false

ggplot(penguins, aes(x = island, fill = species)) +
  geom_bar(alpha = 0.8) +
  scale_fill_manual(values = c("darkorange","purple","cyan4"),
                    guide = "none") +
  theme_minimal() +
  facet_wrap(~species, ncol = 1) +
  coord_flip()

```

Talk more about it.

And also planes (@fig-planes). (You can change the height and width, but don't worry about doing that until you have finished every other aspect of the paper - Quarto will try to make it look nice and the defaults usually work well once you have enough text.)

Talk way more about it.

## Predictor variables

Add graphs, tables and text.

Use sub-sub-headings for each outcome variable and feel free to combine a few into one if they go together naturally.

>>>>

library(tidyverse)
library(palmerpenguins)
library(here)
library(tibble)
library(knitr)
```

@@

For the data we used in this analysis about the polling result for Kamala Harris and Donalad Trump in 2024 USA president election. \
- **response variable:** pct(pct: The percentage of the vote or support that the candidate received in the poll) \
- **numeric predictor:**\
 sample size(sample_size: The total number of respondents participating in the poll) \
 timegap(the time gap between the poll start date and the real election date i.e timegap = real US election date - poll start date) \
 pollscore(A numeric value representing the score or reliability of the pollster in question) \
- **categorical predictor**
state(The U.S. state where the poll was conducted or focused) \
methodology(The method used to conduct the poll)\

@@

In this dataset, each row represents a polling question that records the variables of interest. Each entry allows us to explore the real-world relationships between polling factors and the support percentage (`pct`) for the candidates Kamala Harris and Donald Trump. This dataset enables an analysis of how various polling characteristics influence the reported support levels for the candidates we are focused.

## Clean Data
The data cleaning process involves several steps to ensure the quality and relevance of the polling data. First, we filter the dataset to retain only poll results with a numeric grade of 2.7 or higher, indicating that the polls are considered reliable. Next, we address missing values in the state attribute: polls with NA in the state column are considered national polls.

We then create a new attribute, days_taken_from_election, which represents the time gap between the poll's start date and the actual U.S. election date. Additionally, we filter the dataset to include only polls conducted after July 21, 2024, the date when Kamala Harris declared her candidacy. Finally, we remove any remaining rows that contain missing values to ensure a clean dataset.

```{r}
#| label: tbl-cleaned-data
#| tbl-cap: Sample of cleaned US election data 
#| echo: false
#| warning: false
cleaned_data = read_csv(here("data/02-analysis_data/cleaned_data.csv"))

cleaned_data |>
  select(pct, sample_size, pollscore, days_taken_from_election, state, methodology, answer) |>
  head(6) |>
  kable(
    col.names = c("pct", "sample_size", "pollscore", "days_taken_from_election","state", "methodology", "answer"),
    booktabs = TRUE
    ) 
```

## Basic Statistics Summary for Data

```{r}
#| label: fig-clean-data-histogram-Harris-Trump
#| fig-cap: the average PCT vs State for Harris and Trump
#| echo: false
#| warning: false
#| fig-subcap: ["average PCT vs State for Harris","PCT vs State for Trump"]
#| layout-nrow: 2
#| fig-width: 40
#| fig-height: 20
Harris_data = cleaned_data[cleaned_data$answer == "Harris", ]
Harris_data |>
  group_by(state) |>
  summarise(avg_pct = mean(pct, na.rm = TRUE)) |>
  ggplot(mapping = aes(x = state, y = avg_pct)) + 
  geom_col()  +
  theme_minimal() +
  labs(
    x = "State", y = "support percentage(pct)", title = "average support percentage(pct) for Harris per state"
  )

Trump_data = cleaned_data[cleaned_data$answer == "Trump", ]
Trump_data |>
  group_by(state) |>
  summarise(avg_pct = mean(pct, na.rm = TRUE)) |>
  ggplot(mapping = aes(x = state, y = avg_pct)) + 
  geom_col()  +
  theme_minimal() +
  labs(
    x = "State", y = "support percentage(pct)", title = "average support percentage(pct) for Harris per state"
  )

# cleaned_data %>%
#   mutate(month = factor(month, levels = c("Jan", "Feb", "Mar", "Apr", "May", "Jun",
#                                           "Jul", "Aug", "Sep", "Oct", "Nov", "Dec"))) %>%
#   ggplot(mapping = aes(x = month, y = total_decedents)) + 
#   geom_bar(stat = "identity") +
#   theme_minimal() +
#   labs(
#     x = "Month", 
#     y = "Total Decedents"
#   )
```
